\textbf{ Interpretable Gaussian Processes for Stellar Light Curves }

### 2020-11-09 tex/first_draft/ms.tex
update tex

%
\section{The Size Integrals: \lowercase{$\mathbf{e_1}$} and $\mathbf{E_1}$}
\label{sec:size}
%
\subsection{Functional form}
%

\subsubsection{$l_{\mathrm{max}} = \infty$}
\label{sec:size-function}
%
We adopt the following expression for the $n^{\mathrm{th}}$ term in $\mathbf{s}$
in the expansion of a dark, unit contrast circular spot at
$\lambda = \phi = 0$:
%
\begin{align}
    \label{eq:sinf}
    s_{n}^\infty(\rho) =
    -\dfrac{1}{2}
    \begin{cases}
        \dfrac{\rho}{1 + \rho}
         & l = m = 0
        \\[1em]
        \dfrac{\rho \left( 2 + \rho \right)}
        {\sqrt{2l + 1} (1 + \rho)^{l + 1}}
         & l > 0, \, m = 0
        \\[1em]
        0
         & \mathrm{otherwise}
    \end{cases}
\end{align}
%
where
$l = l(n), m = m(n)$ (Equation~\ref{eq:lm}),
$\delta$ is the Kronecker delta,
$\rho \in [0, \infty)$ is the spot radius,
and the $\infty$ superscript denotes the fact that this expression
is valid in the limit $l_{\mathrm{max}} \rightarrow \infty$
(more on this below).
Equation~(\ref{eq:sinf}) is convenient because it satisfies
five important properties:
%
\begin{enumerate}[itemsep=2pt,parsep=1pt,label=\textbf{\arabic*}]
    \item The surface intensity is azimuthally symmetric about the spot center
    \item The surface intensity monotonically increases away from the spot center
    \item The surface intensity at the spot center is $-1$
    \item The surface intensity at the antipode of the spot center is zero
    \item The size of the spot increases monotonically with $\rho$
\end{enumerate}
%
These properties may be demonstrated by considering the
expression for the surface intensity at a given polar angle $\theta$ and
azimuth $\varphi$:
%
\begin{align}
    \begin{split}
        I(\theta, \varphi)
        & =
        \sum\limits_{n=0}^\infty
        s_{n}^\infty Y_{n}(\theta, \varphi) \\
        & =
        \sum\limits_{n=0}^\infty
        s_{n}^\infty \sqrt{2l + 1} \delta_{m0} P_l(\cos\theta)
    \end{split}
\end{align}
%
where $Y_n$ is the spherical harmonic of degree $l(n)$ and order $m(n)$
(see Equation~\ref{eq:lm}),
$P_l$ is the Legendre polynomial of degree $l$,
and we have implicitly
assumed a normalization such that the integral of our expansion over
the unit sphere is $4\pi$.
Note that the expression above is independent of the azimuth $\varphi$,
a consequence of the fact that all harmonics with $m \ne 0$ are zero; this
expansion is therefore azimuthally symmetric about the spot center, as
stated in \textbf{1}.
%
Combining the above expression with Equation~(\ref{eq:sinf}) and
rearranging, we may write
%
\begin{align}
    \label{eq:Igen}
    I(\theta, \varphi) =
    I(\theta) =
    \dfrac{\rho}{2}
    -
    \dfrac{\rho \left( 2 + \rho \right)}{2 (1 + \rho)}
    \sum\limits_{l=0}^\infty \left(\dfrac{1}{1 + \rho}\right)^l P_l(\cos\theta)
    \quad.
\end{align}
%
The summation in Equation~(\ref{eq:Igen}) has a closed-form expression in
terms of the generating function of the Legendre polynomials:
%
\begin{align}
    \label{eq:gen}
    \sum\limits_{l=0}^\infty t^l P_l(\cos\theta) = \frac{1}{\sqrt{1 + t^2 - 2 t \cos\theta}}
    \quad,
\end{align}
%
so we may express the intensity as a function of polar angle
in the fairly simple form
%
\begin{align}
    \label{eq:Ifinal}
    I(\theta) & = A - \frac{B}{\sqrt{C - \cos\theta}}
    \quad,
    %
    \\
    \intertext{where}
    %
    \begin{split}
        A & = \dfrac{\rho}{2}                      \\
        B & = \rho (2 + \rho) \sqrt{\dfrac{1}{8 + 8 \rho}} \\
        C & = \dfrac{1 + (1 + \rho)^2}{2 + 2 \rho}
    \end{split}
\end{align}
%
are positive constants.

Differentiating Equation~(\ref{eq:Ifinal}) with respect to $\theta$, we have
%
\begin{align}
    \label{eq:Ideriv}
    \dfrac{\mathrm{d}I(\theta)}{\mathrm{d}\theta} & =
    -\frac{B\sin\theta}{2(C - \cos\theta)^\frac{3}{2}}
    \quad,
\end{align}
%
which is zero only for $\theta = 0$ (for which $I(\theta)$ is
minimized) and $\theta = \pi$ (for which it is maximized). The intensity
therefore increases monotonically from the spot center to the antipode,
as stated in \textbf{2}. The value at the minimum is
%
\begin{align}
    \begin{split}
        I_{\mathrm{min}} & = A - \dfrac{B}{\sqrt{C - 1}} \\
        & = -1
        \quad,
    \end{split}
\end{align}
%
as stated in \textbf{3}, and the value at the maximum is
%
\begin{align}
    \begin{split}
        I_{\mathrm{max}} & = A - \dfrac{B}{\sqrt{C + 1}} \\
        & = 0
        \quad,
    \end{split}
\end{align}
%
as stated in \textbf{4}.
Finally, to show \textbf{5}, let us compute the half width at half minimum
$\Delta\theta$ of the intensity profile:
%
\begin{align}
    A - \dfrac{B}{\sqrt{C - \cos{\Delta\theta}}} =
    -\dfrac{1}{2}
\end{align}
%
Solving for $\Delta\theta$ yields
%
\begin{align}
    \label{eq:hwhm}
    \Delta\theta =
    \cos^{-1} \left[ \dfrac{2 + 3 \rho (2 + \rho)}{2 (1 + \rho)^3} \right]
    \quad.
\end{align}
%
Differentiation with respect to $r$ yields
%
\begin{align}
    \dfrac{\mathrm{d}\Delta\theta}{\mathrm{d}\rho} =
    \frac{3}{\left(1 + \rho\right)
        \sqrt{\left(1 + 2 \rho\right)
            \left(3 + 2 \rho\right)}}
    \quad,
\end{align}
%
which is positive definite for all $\rho > 0$.

@@

\subsubsection{Finite $l_{\mathrm{max}}$}
%
It is important to note that the
properties of the spot expansion described above
(in particular its monotonicity and values at the spot center and antipode)
are inherited from the generating function of the Legendre polynomials
(Equation~\ref{eq:gen}). These properties are therefore only rigorously
true when the expansion is taken to spherical
harmonic degree $l_{\mathrm{max}} = \infty$. Expansions truncated to finite
$l_{\mathrm{max}}$ will in general experience ringing and a spot amplitude
whose (absolute) value is in general smaller than unity.

To mitigate these issues, we will do two things: we will impose a minimum
radius $r_{\mathrm{min}}$ to prevent highly oscillatory behavior within
the domain of our function; and we will scale the spot expansion
by an overall amplitude factor so that
the peak spot intensity is as close to (negative) unity as possible
everywhere within the domain.
We find that this can be accomplished by expressing the
$n^{\mathrm{th}}$ term in the spot expansion $\mathbf{s}$ as
%
\begin{align}
    s_n(\rho) = s_n^\infty\left( c_0 + c_1\rho \right) \left(1 + c_2 \left(1 - \rho\right) ^ {c_3}\right)
\end{align}
%
where $c_0, c_1, c_2, c_3$ are constants at a given
value of $l_{\mathrm{max}}$.
%
Plugging this into Equation~(\ref{eq:sinf}), we have
%
\begin{align}
    \label{eq:s}
    s_{n}(\rho) =
    -\dfrac{1 + c_2 \left(1 - \rho\right) ^ {c_3}}{2}
    \begin{cases}
        \dfrac{r\left(\rho\right)}{1 + r\left(\rho\right)}
         & l = m = 0
        \\[1em]
        \dfrac{ r\left(\rho\right) \big( 2 + r\left(\rho\right) \big)}
        {\sqrt{2l + 1} \big(1 + r\left(\rho\right)\big)^{l + 1}}
         & l > 0, \, m = 0
        \\[1em]
        0
         & \mathrm{otherwise}
    \end{cases}
    \quad,
\end{align}
%
where we define
%
\begin{align}
    r\left(\rho\right) \equiv c_0 + c_1\rho
    \quad.
\end{align}
%
Note that in the limit $l_{\mathrm{max}} \rightarrow \infty$, we may set
$c_0 = 0$, $c_1 = 1$, and $c_2 = 0$ to recover the original parametrization.
At finite $l_{\mathrm{max}}$, however, we solve for the $c_n$ numerically,
by requiring the following conditions:

>>>>

%
\section{The Size Integrals: \lowercase{$\mathbf{e_1}$} and $\mathbf{E_1}$}
\label{sec:size}
%
\subsection{Functional form}
%
\xxx{TODO: Explain why we need the corrected, scaled radius.}

\subsubsection{$l_{\mathrm{max}} = \infty$}
\label{sec:size-function}
%
We adopt the following expression for the $n^{\mathrm{th}}$ term in $\mathbf{s}$
in the expansion of a dark, unit contrast circular spot at
$\lambda = \phi = 0$:
%
\begin{align}
    \label{eq:sinf}
    s_{n}'(\rho') =
    -\dfrac{1}{2\sqrt{2l + 1}}
    \begin{cases}
        \dfrac{\rho'}{1 + \rho'}
         & l = m = 0
        \\[1em]
        \dfrac{\rho' \left( 2 + \rho' \right)}
        {(1 + \rho')^{l + 1}}
         & l > 0, \, m = 0
        \\[1em]
        0
         & \mathrm{otherwise}
    \end{cases}
\end{align}
%
where
$l = l(n), m = m(n)$ (Equation~\ref{eq:lm}),
$\delta$ is the Kronecker delta,
$\rho' \in [0, \infty)$ is the spot radius,
and the prime superscript on both $s_{n}'$ and $\rho'$ denotes
the fact that this expression
is valid only in the limit $l_{\mathrm{max}} \rightarrow \infty$
(more on this below).
Equation~(\ref{eq:sinf}) is convenient because it satisfies
five important properties:
%
\begin{enumerate}[itemsep=2pt,parsep=1pt,label=\textbf{\arabic*}]
    \item The surface intensity is azimuthally symmetric about the spot center
    \item The surface intensity monotonically increases away from the spot center
    \item The surface intensity at the spot center is $-1$
    \item The surface intensity at the antipode of the spot center is zero
    \item The size of the spot increases monotonically with $\rho'$
\end{enumerate}
%
These properties may be demonstrated by considering the
expression for the surface intensity at a given polar angle $\theta$ and
azimuth $\varphi$:
%
\begin{align}
    \begin{split}
        I(\theta, \varphi)
        & =
        \sum\limits_{n=0}^\infty
        s_{n}' Y_{n}(\theta, \varphi) \\
        & =
        \sum\limits_{n=0}^\infty
        s_{n}' \sqrt{2l + 1} \delta_{m0} P_l(\cos\theta)
    \end{split}
\end{align}
%
where $Y_n$ is the spherical harmonic of degree $l(n)$ and order $m(n)$
(see Equation~\ref{eq:lm}),
$P_l$ is the Legendre polynomial of degree $l$,
and we have implicitly
assumed a normalization such that the integral of our expansion over
the unit sphere is $4\pi$.
Note that the expression above is independent of the azimuth $\varphi$,
a consequence of the fact that all harmonics with $m \ne 0$ are zero; this
expansion is therefore azimuthally symmetric about the spot center, as
stated in \textbf{1}.
%
Combining the above expression with Equation~(\ref{eq:sinf}) and
rearranging, we may write
%
\begin{align}
    \label{eq:Igen}
    I(\theta, \varphi) =
    I(\theta) =
    \dfrac{\rho'}{2}
    -
    \dfrac{\rho' \left( 2 + \rho' \right)}{2 (1 + \rho')}
    \sum\limits_{l=0}^\infty \left(\dfrac{1}{1 + \rho'}\right)^l P_l(\cos\theta)
    \quad.
\end{align}
%
The summation in Equation~(\ref{eq:Igen}) has a closed-form expression in
terms of the generating function of the Legendre polynomials:
%
\begin{align}
    \label{eq:gen}
    \sum\limits_{l=0}^\infty t^l P_l(\cos\theta) = \frac{1}{\sqrt{1 + t^2 - 2 t \cos\theta}}
    \quad,
\end{align}
%
so we may express the intensity as a function of polar angle
in the fairly simple form
%
\begin{align}
    \label{eq:Ifinal}
    I(\theta) & = A - \frac{B}{\sqrt{C - \cos\theta}}
    \quad,
    %
    \\
    \intertext{where}
    %
    \begin{split}
        A & = \dfrac{\rho'}{2}                      \\
        B & = \rho' (2 + \rho') \sqrt{\dfrac{1}{8 + 8 \rho'}} \\
        C & = \dfrac{1 + (1 + \rho')^2}{2 + 2 \rho'}
    \end{split}
\end{align}
%
are positive constants.

Differentiating Equation~(\ref{eq:Ifinal}) with respect to $\theta$, we have
%
\begin{align}
    \label{eq:Ideriv}
    \dfrac{\mathrm{d}I(\theta)}{\mathrm{d}\theta} & =
    -\frac{B\sin\theta}{2(C - \cos\theta)^\frac{3}{2}}
    \quad,
\end{align}
%
which is zero only for $\theta = 0$ (for which $I(\theta)$ is
minimized) and $\theta = \pi$ (for which it is maximized). The intensity
therefore increases monotonically from the spot center to the antipode,
as stated in \textbf{2}. The value at the minimum is
%
\begin{align}
    \begin{split}
        I_{\mathrm{min}} & = A - \dfrac{B}{\sqrt{C - 1}} \\
        & = -1
        \quad,
    \end{split}
\end{align}
%
as stated in \textbf{3}, and the value at the maximum is
%
\begin{align}
    \begin{split}
        I_{\mathrm{max}} & = A - \dfrac{B}{\sqrt{C + 1}} \\
        & = 0
        \quad,
    \end{split}
\end{align}
%
as stated in \textbf{4}.
Finally, to show \textbf{5}, let us compute the half width at half minimum
$\Delta\theta$ of the intensity profile:
%
\begin{align}
    A - \dfrac{B}{\sqrt{C - \cos{\Delta\theta}}} =
    -\dfrac{1}{2}
\end{align}
%
Solving for $\Delta\theta$ yields
%
\begin{align}
    \label{eq:hwhm}
    \Delta\theta(\rho') =
    \cos^{-1} \left[ \dfrac{2 + 3 \rho' (2 + \rho')}{2 (1 + \rho')^3} \right]
    \quad.
\end{align}
%
Differentiation with respect to $r$ yields
%
\begin{align}
    \dfrac{\mathrm{d}\Delta\theta}{\mathrm{d}\rho'} =
    \frac{3}{\left(1 + \rho'\right)
        \sqrt{\left(1 + 2 \rho'\right)
            \left(3 + 2 \rho'\right)}}
    \quad,
\end{align}
%
which is positive definite for all $\rho' > 0$.

For future reference, Equation~(\ref{eq:hwhm}) may be inverted to obtain the
radius $\rho'$ corresponding to a particular value of the half width
at half minimum:
%
\begin{align}
    \label{eq:hwhm}
    \rho'(\Delta\theta) =
    \left(\dfrac{1 +
        \cos\left(\dfrac{2\Delta\theta}{3}\right) +
        \sqrt{3}\sin\left(\dfrac{2\Delta\theta}{3}\right)}
    {2 \cos\Delta\theta}\right) - 1
    \quad.
\end{align}
%

@@

\subsubsection{Finite $l_{\mathrm{max}}$}
%
The
properties of the spot expansion described above
(in particular its monotonicity and values at the spot center and antipode)
are inherited from the generating function of the Legendre polynomials
(Equation~\ref{eq:gen}). These properties are therefore only rigorously
true when the expansion is taken to spherical
harmonic degree $l_{\mathrm{max}} = \infty$. Expansions truncated to finite
$l_{\mathrm{max}}$ will in general experience ringing and a spot amplitude
whose (absolute) value is in general smaller than unity.

To mitigate these issues, we will do two things: we will impose a minimum
radius $r_{\mathrm{min}}$ to prevent highly oscillatory behavior within
the domain of our function; and we will scale the spot expansion
by an overall amplitude factor so that
the peak spot intensity is as close to (negative) unity as possible
everywhere within the domain.
We find that this can be accomplished by expressing the
$n^{\mathrm{th}}$ term in the spot expansion $\mathbf{s}$ as
%
\begin{align}
    s_n(\rho) = s_n'\left( \rho' \right) \left(1 + c_2 \left(1 - \rho\right) ^ {c_3}\right)
\end{align}
%
where
%
\begin{align}
    \begin{split}
        \rho' &= \rho'\left(\rho\right) \\ &\equiv c_0 + c_1\rho
        \quad.
    \end{split}
\end{align}
%
and $c_0, c_1, c_2, c_3$ are constants at a given
value of $l_{\mathrm{max}}$.
%
Plugging this into Equation~(\ref{eq:sinf}), we have
%
\begin{align}
    \label{eq:s}
    s_{n}(\rho) =
    -\dfrac{1 + c_2 \left(1 - \rho\right) ^ {c_3}}{2\sqrt{2l + 1}}
    \begin{cases}
        \dfrac{\rho'}{1 + \rho'}
         & l = m = 0
        \\[1em]
        \dfrac{ \rho' \left( 2 + \rho' \right)}
        {\left(1 + \rho'\right)^{l + 1}}
         & l > 0, \, m = 0
        \\[1em]
        0
         & \mathrm{otherwise}
    \end{cases}
    \quad.
\end{align}
%
Note that in the limit $l_{\mathrm{max}} \rightarrow \infty$, we may set
$c_0 = 0$, $c_1 = 1$, and $c_2 = 0$ to recover the original parametrization.
At finite $l_{\mathrm{max}}$, however, we solve for the $c_n$ numerically,
by requiring the following conditions:
